---
title: "strings_and_factors"
author: "Matthew Lawlor"
date: "10/20/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

library(p8105.datasets)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
  
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Strings and regex

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")

str_replace(string_vec, "jeff", "Jeff")
```

For exact matches, you can designate matches at the beginning or end of a line.

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think") ##detect string at start of line

str_detect(string_vec, "i think$") ##detect string at end of line
```

You can designate a list of characters that will count as a match.

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush") ##identify upper or lower case of characters in ()
```

You don’t have to list these; instead, you can provide a range of letters or numbers that count as a match.

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

The character . matches anything.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") ##if need actual ., use 7\\.11
```

Some characters are “special”. These include [ and ], ( and ), and .. If you want to search for these, you have to indicate they’re special using \. Unfortunately, \ is also special, so things get weird.

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## Factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))

as.numeric(vec_sex)
```

What happens if i relevel

```{r}
vec_sex = fct_relevel(vec_sex, "male")

as.numeric(vec_sex)
```

## NSDUH

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)

tabl_marj = 
drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>% 
  slice(-1) %>% 
  as_tibble()
```

```{r}
data_marj =
  tabl_marj %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  ) %>% 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

## NSDUH -- factors

```{r}
data_marj %>% 
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% ##note using fct_*** function also recodes vector from character to factor variable
  ggplot(aes(State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Reorder or relevel factors for plots

```{r}
weather_df %>% 
  mutate(name = fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

Reorder or relevel factors for regression modeling (ie change the reference)

```{r}
weather_df %>% 
  mutate(name = fct_reorder(name, tmax)) %>% 
  lm(tmax ~ name, data = .)
```

